Anomalous waves triggered by abrupt depth changes: laboratory experiments and truncated KdV statistical mechanics

### 2019-11-20 paper2_JNLS/paper.tex
NM

\begin{abstract} 
Recent laboratory experiments revealed the emergence of anomalous statistics in a randomized surface waves upon encountering an abrupt depth change (ADC) in the bottom topography (Bolles et al. PRF 2019). Majda et al.~2019 established a theoretical foundation, based on analysis of the truncated KdV (TKdV) system, that successfully captures several key qualitative features of the experiments including the robust emergence of anomalous statistics. The framework exploits the Hamiltonian structure of TKdV to construct invariant measures that are statistical matched at the ADC, and is corroborated by simulations of the deterministic dynamics. 
%
Here we present a more comprehensive treatment of these parallel experimental and modeling efforts, with emphasis on the link between the two. We present an assortment of new experimental measurements and carefully compare them to model predictions. We establish a precise link between experimental and model parameters, so that the invariant measure underlying the model can be calibrated to the specific conditions present in experiments. The model predicts a peculiar relationship between surface-displacement skewness and surface-slope variance, which is confirmed in a spectacular fashion by direct experimental measurements.
\end{abstract}

@@

A recent series of laboratory \cite{bolles2019} and numerical investigations \cite{viotti2014} have demonstrated the emergence of anomalous wave statistics from abrupt variations in bottom topography. Since topographical variations are strictly one-dimensional, these studies can be viewed as offering a bare minimum set of conditions capable of generating anomalous waves. In particular, the more complex mechanism of focusing by 2D bottom topography is  absent. The studies thus offer a paradigm system, with emergent anomalous statistical properties similar to that seen in more complex systems, but in a tractable context that is amenable to analysis.

>>>>

\begin{abstract} 
Recent laboratory experiments of \boetal revealed the emergence of anomalous statistics in a randomized wave-field that encounters an abrupt change in bottom topography. \maetal established a theoretical foundation, based on analysis of the truncated Korteweg–de Vries (TKdV) system, that successfully captures several key qualitative features of the experiments, including the robust emergence of anomalous statistics in the outgoing waves. The framework exploits the Hamiltonian structure of TKdV to construct invariant measures that are statistical matched at the depth change, and is corroborated by numerical simulations of the deterministic dynamics. 
Here we present a more comprehensive treatment of these parallel experimental and modeling efforts, with emphasis on the link between the two. We present an assortment of new experimental measurements and carefully compare them to model predictions. We establish a precise link between experimental and model parameters, so that the invariant measure underlying the model can be calibrated to the specific conditions present in experiments and a quantitative comparison can be made. The model predicts a peculiar relationship between surface-displacement skewness and surface-slope variance, which is confirmed in a spectacular fashion by new experimental measurements.
\end{abstract}

@@

A recent series of laboratory \cite{bolles2019} and numerical investigations \cite{viotti2014} have demonstrated the emergence of anomalous wave statistics from abrupt variations in bottom topography. Since topographical variations are strictly one-dimensional, these studies can be viewed as offering a bare minimum set of conditions capable of generating anomalous waves. In particular, the more complex mechanism of focusing by 2D bottom topography is  absent. The studies thus offer a paradigm system, with emergent anomalous features similar to those seen in more complex systems, but in a tractable context that is amenable to analysis.
